Generative AI for image reconstruction\\ in Intensity Interferometry: a first attempt

### 2025-07-17 Phase_II/draft/GAN.tex
figure
..
edits

\section{Generative Adversarial Networks}
\begin{figure*}[hbt]
	%\centering
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}\hfil
	%\caption{Discriminator loss for three different learning rates. }
	%	\label{fig:Plot_learning_rate_discloss}
	%\end{subfigure}\hfill
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
	%	\label{fig:Plot_learning_rate_genloss}
	%\end{subfigure}\hfill
	\caption{Generator and discriminator losses for three different learning rates. The left figure shows the total discriminator loss, and the right figure shows the total generator loss. There is no significant difference, but the highest learning rate might be prone to outliers. }
	\label{fig:Plot_learning_rate_loss}
\end{figure*}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

\begin{figure*}[hbt]

@@

  \caption{Generator and discriminator losses for three different kernel sizes in the convolutional layers. Here, the smallest kernel size has many outliers, while the largest kernel size seems to be the most stable.}

@@

  \caption{The amount of Discriminator training has a high impact on the Discriminator score. The Discriminator repetition indicates the factor by which the Discriminator is trained more with respect to the Generator.}

@@

  \includegraphics[width=\textwidth]{fig/analysis/Plot_N_telescopes_gen_total_loss.pdf}\hfil

@@

  \caption{The number of telescopes has a very impact on the model performance. If there are only two telescopes (1 baseline), both Discriminator and Generator are not trained smoothly: One gains a big advantage over the other. The result of four telescopes (6 baselines) is a lot better because the loss functions change only slightly with increasing steps.}

>>>>

\section{Generative Adversarial Networks}
\begin{figure*}
	%\centering
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}\hfil
	%\caption{Discriminator loss for three different learning rates. }
	%	\label{fig:Plot_learning_rate_discloss}
	%\end{subfigure}\hfill
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		%\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
	%	\label{fig:Plot_learning_rate_genloss}
	%\end{subfigure}\hfill
	\caption{Discriminator and Generator losses for three different learning rates. The left panel shows the total discriminator loss, and the right figure shows the total generator loss (Eq.~\ref{eq:total_gen_loss}).  There is no significant difference, but the highest learning rate might be prone to outliers.}
	\label{fig:Plot_learning_rate_loss}
\end{figure*}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

\begin{figure*}

@@

  \caption{Generator and discriminator losses (left and right panels respectively) for three different kernel sizes in the convolutional layers. Here, the smallest kernel size has many outliers, while the largest kernel size seems to be the most stable.}

@@

  \caption{The left panel shows the Discriminator loss for different Discriminator training, while the right panel shows the Generator losses for different Discriminator training.  The amount of Discriminator training has a high impact on the Discriminator score. The Discriminator repetition indicates the factor by which the Discriminator is trained more with respect to the Generator.}

@@

  \includegraphics[width=0.49\textwidth]{fig/analysis/Plot_N_telescopes_gen_total_loss.pdf}\hfil

@@

  \caption{Discriminator loss (left panel) and Generator loss (right panel) for different numbers of telescopes.  The number of telescopes has a very impact on the model performance. If there are only two telescopes (1 baseline), both Discriminator and Generator are not trained smoothly: One gains a big advantage over the other. The result of four telescopes (6 baselines) is a lot better because the loss functions change only slightly with increasing steps.}

### 2025-07-17 Phase_II/draft/II.tex
..
edits

	\centering

@@

	\centering

@@

	\caption{The left panel shows the absolute value of the two-dimensional Fast Fourier Transform of fig.~\ref{fig:image} measured by baselines shown in fig.~\ref{fig:teles}. The right panel shows the same on logarithmic scale}

@@

where $\tau$ is the time delay between the telescopes. For spatially coherent and randomly polarized light, eq.~(\ref{eqn:HBT}) reduces to the relation \citep[sometimes called the Siegert relation, see e.g.,][]{acciari2020optical}.

@@

where $\Delta f$ is the electronic bandwidth of the photon detectors which measure the intensities.  Values of $\Delta\nu\sim 1\,\mathrm{THz}$ and $\Delta f \sim 1\,\mathrm{GHz}$ are typical of recent work.  In eq.~(\ref{eqn:HBT2}), $V_{12}$, referred to as the complex visibility function, is the Fourier transform of the source brightness distribution. For a uniform disk it is given by

@@

where $\theta_D$ is the angular diameter of the star and $b$ is the radial coordinate in the conventional interferometric $(u , v) = (x/\lambda, y/\lambda)$ plane, with $\lambda$ representing the optical wavelength of the filter used for observation. It is evident from eq.~(\ref{eqn:absvisib}) that $V_{12}$ contains information about the star's angular diameter. However, the phase information is lost since we measure only the absolute value $\vert V_{12} \vert^2$. In observational astronomy, the correlation is often expressed in terms of the normalized contrast, given by:

@@

Intensity Interferometry counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. Fig.\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on both linear and logarithmic scales. In practice, only a small part of this information is available, because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. Therefore, we have simulated the II observation of a sky source by four telescopes (correlated in pairs) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.
\begin{figure}[hbt]
	\centering
	\includegraphics[width=\linewidth]{fig/ellipse1612.jpg}
	\caption{Merged image, which includes the original and the sparsely sampled Fourier plane. It is exactly what the GAN receives. }
	\label{fig:GANinput}
\end{figure}

>>>>

\begin{figure}

@@

@@

  \caption{The left panel shows the absolute value of the two-dimensional Fast Fourier Transform of Fig.~\ref{fig:image} measured by baselines shown in Fig.~\ref{fig:teles}. The right panel shows the same on logarithmic scale}

@@

where $\tau$ is the time delay between the telescopes. For spatially coherent and randomly polarized light, Eq.~\eqref{eqn:HBT} reduces to the relation \citep[sometimes called the Siegert relation, see e.g.,][]{acciari2020optical}.

@@

where $\Delta f$ is the electronic bandwidth of the photon detectors which measure the intensities.  Values of $\Delta\nu\sim 1\,\mathrm{THz}$ and $\Delta f \sim 1\,\mathrm{GHz}$ are typical of recent work.  In Eq.~\eqref{eqn:HBT2}, $V_{12}$, referred to as the complex visibility function, is the Fourier transform of the source brightness distribution. For a uniform disk it is given by

@@

where $\theta_D$ is the angular diameter of the star and $b$ is the radial coordinate in the conventional interferometric $(u , v) = (x/\lambda, y/\lambda)$ plane, with $\lambda$ representing the optical wavelength of the filter used for observation. It is evident from Eq.~\eqref{eqn:absvisib} that $V_{12}$ contains information about the star's angular diameter. However, the phase information is lost since we measure only the absolute value $\vert V_{12} \vert^2$. In observational astronomy, the correlation is often expressed in terms of the normalized contrast, given by:

@@

Intensity Interferometry counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. Fig.~\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on both linear and logarithmic scales. In practice, only a small part of this information is available, because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. Therefore, we have simulated the II observation of a sky source by four telescopes (correlated in pairs) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.
\begin{figure}
  \includegraphics[width=\linewidth]{fig/ellipse1612.jpg}
  \caption{Merged image, which includes the original and the sparsely sampled Fourier plane. It is exactly what the GAN receives.}
  \label{fig:GANinput}
\end{figure}
